---
title: "numeric eda"
author: "Binyue Hu"
date: "2022-10-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's import data

```{r}
library(tidyverse)
```

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group_by()`

```{r}
weather_df %>%
  group_by(name,month)
```

## `summarize()`

let's group and then summarize!

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n())
```

can we count in other ways?

```{r}
weather_df %>%
  count(name)
```

!!! don't use `table()` 

```{r eval = FALSE}
weather_df %>%
  pull(month) %>% 
  table
```

you can have multiple summaries, we use `distinct()` to count the number of distinct values 

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date)
    )
```


## a brief aside on 2x2 tables 

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

the last line 'TRUE' in case_when function is the default thing to make sure 
every value in the data frame is counted and not be left. 
```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

## General Summaries 

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

