---
title: "viz-part-01"
author: "Binyue Hu"
date: "2022-10-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's import data

```{r}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Let's make a scatterplot 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

let's make the same scatterplot different 

```{r}
weather_df %>%
  drop_na() %>%
  filter(name == "CentralPark_NY") %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Let's keep making the same plot but different 

```{r}

plot_weather = 
  weather_df %>%
  drop_na() %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()

```


## Let's fancy this up a bit

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

ä¸Šé¢è¿™æ®µcodeå’Œä¸‹é¢è¿™æ®µcodeæ˜¯åŒä¸€ä¸ªæ„æ€,ä¸‹é¢è¿™æ®µcodeé‡Œçš„è¿™ä¸ª`smooth`å‡½æ•°è®©ggploté‡Œå¤šäº†ä¸€æ¡çº¿æ¥ä½“çŽ°æ•´ä½“çš„è¶‹åŠ¿

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth()
```

ðŸ‘‡ This code remove `color = name`from `geom_point`to `ggplot`, it causes that different names with different colors have different smooth lines. 
`alpha ` changes the transparancy of points 
`se = FALSE` is get rid of error bar 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)
```

use `facet_grid` to make separate panels 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
   facet_grid(. ~ name)
# left side of "~" defines rows and right side of "~" defines column 
```

`tmax` vs ` tmin` is boring, let's spice it up some. 
we change the x-aix into date 

```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```

we change the color equal to percipitation 
```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = prcp)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```

we keep both color equal to name and use size of point to show the percipitation 
```{r}
 weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## some quick stuff

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```


